### 2010-04-27 paper/paper.tex
Add paper outline

\title{ Data Encryption Standard Implementation 4003-482-01 RIT }

\author{
Brian Gianforcaro \and Sam Milton \\
}

\date{\today}

\documentclass[12pt]{article}

\usepackage{graphicx}
\usepackage{multicol}
\usepackage{mdwlist}
\usepackage{amssymb,amsmath}
\usepackage{cite}
\usepackage[bookmarks=false,colorlinks=true,linkcolor={blue},pdfstartview={XYZ null null 1.22}]{hyperref}
\usepackage{url}
\usepackage[T1]{fontenc}

\begin{document}
\maketitle

\begin{abstract}
DES (Data Encryption Standard) is the 56-bit block cipher that was once used to encrypt all US Goverment non-classified documents. 
\end{abstract}

\newpage      
\section{The Cryptographic Primative}

\section{Examples}
\section{Software Design}
\section{Original Runtime Measurements}
\section{Original Runtime Analysis}
\section{Revised Software design}
\section{Revised Runtime Measurements}
\section{Revised Runtime Analysis}
\section{Developers manual}
\section{Users Manual}
\section{ What we learned }
\section{Future Work}
\section{Team Member Work}
\section{References}
\begin{thebibliography}{99}
\bibitem{python}{\href{ http://python.org }{ http://python.org }}
\bibitem{wiki} {\href{ http://en.wikipedia.org/wiki/Latitude_and_longitude_of_cities }{ http://en.wikipedia.org/wiki/Latitude\_and\_longitude\_of\_cities }}
\bibitem{tsp} Montgomery, John {\it Tackling The Traveling Salesman Problem }{\href{ http://www.psychicorigami.com/category/tsp/ }{ http://www.psychicorigami.com/category/tsp/ } }, {\bf 2007}
\bibitem{google} {\href{ http://code.google.com/apis/maps/documentation/staticmaps/  }{ http://code.google.com/apis/maps/documentation/staticmaps/ }}
\bibitem{sa}{\href{ http://www.svengato.com/salesman.html }{ http://www.svengato.com/salesman.html }}
\bibitem{dis}{\href{ http://www.geesblog.com/2009/01/calculating-distance-between-latitude-longitude-pairs-in-python/ }{ http://www.geesblog.com/2009/01/calculating-distance-between-latitude-longitude-pairs-in-python/ }}
\bibitem{LHK}{\href{ http://www.akira.ruc.dk/~keld/research/LKH/ }{ http://www.akira.ruc.dk/~keld/research/LKH/ }}
\end{thebibliography}

\end{document}

### 2010-04-27 src/encrypt.cpp

@@ -17,8 +17,8 @@ int main( ) {
   char key[17] = "01334567";
   char msg[17] = "01234567";
 
-  printf( "Encrypting: %s\n", msg );
-  printf( "With Key: %s\n", key );
+  //printf( "Encrypting: %s\n", msg );
+  //printf( "With Key: %s\n", key );
 
   uint8_t kblck[64];
   uint8_t mblck[64];
@@ -30,17 +30,18 @@ int main( ) {
     0, 0, 0, 0, 0, 0, 0, '\0'
   };
 
-  for ( int i = 0; i < 1; i++ ) {
+  for ( int i = 0; i < 3000000; i++ ) {
     DES cipher( mblck, kblck );
     cipher.encrypt();
     DES::blktostr( cipher.cipherText(), dest );
   }
 
-  printf( "Cipher: ");
+  /*printf( "Cipher: ");
 
   for( int i = 0; i < 7; i++ ) {
     printf( "%c",(unsigned char) dest[i] );
   }
+  */
 
   return 0;
 }

### 2010-05-01 src/blockmode.cpp
Introduce block mode and fix merge of encrypt.

@@ -1,0 +1,27 @@
+/*
+**  File: blockmode.cpp
+**
+**  Authors:
+**
+**     Sam Milton        (srm2997@cs.rit.edu)
+**     Brian Gianforcaro (bjg1955@cs.rit.edu)
+**
+**  Description:
+**
+**  Implementation of all the block modes for DES.
+*/
+
+#include "blockmode.hpp"
+
+BLOCKMODE::BLOCKMODE( mode_t mode ) {
+
+}
+
+BLOCKMODE::~BLOCKMODE( void ) {
+
+}
+
+
+BLOCKMODE::encrypt( char file[], char key[] ) {
+
+}

### 2010-05-01 src/blockmode.hpp

@@ -1,0 +1,28 @@
+/*
+**  File: blockmode.hpp
+**
+**  Authors: 
+** 
+**     Sam Milton        (srm2997@cs.rit.edu)
+**     Brian Gianforcaro (bjg1955@cs.rit.edu)
+**
+*/
+
+enum mode_t {
+  CBC, ECB
+};
+
+
+class BLOCKMODE {
+
+  private:
+
+  public:
+
+    BLOCKMODE( mode_t mode );
+
+    ~BLOCKMODE( void );
+
+    char* encrypt( char file[], char key[] );
+
+}

### 2010-05-01 src/encrypt.cpp

@@ -14,34 +14,16 @@
 
 int main( ) {
 
-  char key[17] = "01334567";
-  char msg[17] = "01234567";
+  uint8_t msg[64] = { 0,0,0,0,0,0,0,1,0,0,1,0,0,0,1,1,0,1,0,0,0,1,0,1,
+                      0,1,1,0,0,1,1,1,1,0,0,0,1,0,0,1,1,0,1,0,1,0,1,1,
+                      1,1,0,0,1,1,0,1,1,1,1,0,1,1,1,1 };
 
-  //printf( "Encrypting: %s\n", msg );
-  //printf( "With Key: %s\n", key );
+  uint8_t key[64] = { 0,0,0,1,0,0,1,1,0,0,1,1,0,1,0,0,0,1,0,1,0,1,1,1,0,1,1,1,1,0,0,1,1,0,0,1,1,0,1,1,1,0,1,1,1,1,0,0,1,1,0,1,1,1,1,1,1,1,1,1,0,0,0,1};
 
-  uint8_t kblck[64];
-  uint8_t mblck[64];
-
-  DES::sttoblk( kblck, key );
-  DES::sttoblk( mblck, msg );
-
-  char dest[8] = {
-    0, 0, 0, 0, 0, 0, 0, '\0'
-  };
-
-  for ( int i = 0; i < 3000000; i++ ) {
-    DES cipher( mblck, kblck );
+  for ( int i = 0; i < 1000000; i++ ) {
+    DES cipher( msg, key );
     cipher.encrypt();
-    DES::blktostr( cipher.cipherText(), dest );
   }
-
-  /*printf( "Cipher: ");
-
-  for( int i = 0; i < 7; i++ ) {
-    printf( "%c",(unsigned char) dest[i] );
-  }
-  */
 
   return 0;
 }

### 2010-05-01 src/blockmode.cpp
Initial version of blockmode

@@ -10,10 +10,15 @@
 **
 **  Implementation of all the block modes for DES.
 */
+#include <iostream>
+#include <fstream>
+
+using namespace std;
 
 #include "blockmode.hpp"
+#include "des.hpp"
 
-BLOCKMODE::BLOCKMODE( mode_t mode ) {
+BLOCKMODE::BLOCKMODE( ) {
 
 }
 
@@ -22,6 +27,138 @@ BLOCKMODE::~BLOCKMODE( void ) {
 }
 
 
-BLOCKMODE::encrypt( char file[], char key[] ) {
+void BLOCKMODE::encrypt( char file[], char outfile[], char key[] ) {
+
+  const size_t BUFFSIZE = 8;
+  char buffer[BUFFSIZE];
+  char obuffer[BUFFSIZE];
+
+  uint8_t keyblock[BKSIZE];
+
+  DES::sttoblk( keyblock, key );
+
+
+  ifstream infile;
+  infile.open( file, ios::binary );
+
+  ofstream ofile;
+  ofile.open( outfile, ofstream::binary );
+
+
+  int padding;
+  uint8_t inblock[BKSIZE];
+  printf("Got to while");
+  while( (padding = 8 - infile.readsome( buffer, BUFFSIZE ) ) == 0 ){
+    printf("\n");
+
+    DES::sttoblk( inblock, buffer );
+
+    DES cipher( inblock, keyblock );
+
+    cipher.encrypt();
+
+    uint8_t* out = cipher.cipherText();
+
+    DES::blktostr( out, obuffer );
+
+    for (int i = 0; i < 8; i++ ) {
+      printf("%c",obuffer[i]);
+
+    }
+    ofile.write( obuffer, BUFFSIZE );
+
+  }
+
+  printf("\n");
+
+  for (int i = 0; i < padding; i++ ) {
+    if ( i+1 == padding ) {
+      buffer[7-i] = 1;
+    } else {
+      buffer[7-i] = 0;
+    }
+  }
+  DES::sttoblk( inblock, buffer );
+  DES cipher( inblock, keyblock );
+  cipher.encrypt();
+  uint8_t* out = cipher.cipherText();
+  DES::blktostr( out, obuffer );
+  for (int i = 0; i < 8; i++ ) {
+    printf("%c",obuffer[i]);
+
+  }
+  ofile.write( obuffer, BUFFSIZE );
+
+  infile.close();
+  ofile.close();
 
 }
+
+
[…]

### 2010-05-01 src/des.cpp

@@ -34,7 +34,7 @@ using namespace std;
 **
 */
 
-DES::DES( block_t block , block_t key ) {
+DES::DES( uint8_t* block , uint8_t* key ) {
 
   assert( block != NULL );
   assert( key != NULL );

### 2010-05-01 src/encrypt.cpp

@@ -11,19 +11,36 @@
 #include <cstdio>
 #include <cstring>
 #include "des.hpp"
+#include "blockmode.hpp"
 
-int main( ) {
 
-  uint8_t msg[64] = { 0,0,0,0,0,0,0,1,0,0,1,0,0,0,1,1,0,1,0,0,0,1,0,1,
+void usage( char file[] ) {
+  printf( "Usage: %s <infile> <outfile> key\n ", file );
+}
+
+int main( int argc, char* argv[] ) {
+
+  if ( argc < 4 || argc > 4 ) {
+    usage( argv[0] );
+  }
+
+
+    /* uint8_t msg[64] = { 0,0,0,0,0,0,0,1,0,0,1,0,0,0,1,1,0,1,0,0,0,1,0,1,
                       0,1,1,0,0,1,1,1,1,0,0,0,1,0,0,1,1,0,1,0,1,0,1,1,
                       1,1,0,0,1,1,0,1,1,1,1,0,1,1,1,1 };
 
   uint8_t key[64] = { 0,0,0,1,0,0,1,1,0,0,1,1,0,1,0,0,0,1,0,1,0,1,1,1,0,1,1,1,1,0,0,1,1,0,0,1,1,0,1,1,1,0,1,1,1,1,0,0,1,1,0,1,1,1,1,1,1,1,1,1,0,0,0,1};
+  */
 
+  char* key = argv[3];
+  BLOCKMODE encrypter;
+  encrypter.encrypt( argv[1], argv[2], key );
+  /*
   for ( int i = 0; i < 1000000; i++ ) {
     DES cipher( msg, key );
     cipher.encrypt();
   }
+  */
 
   return 0;
 }

### 2010-05-01 src/blockmode.cpp
Add some memset's to blockmode

@@ -10,6 +10,12 @@
 **
 **  Implementation of all the block modes for DES.
 */
+
+#include <cstdio>
+#include <cstring>
+#include <cstdlib>
+#include <cassert>
+#include <stdint.h>
 #include <iostream>
 #include <fstream>
 
@@ -34,6 +40,7 @@ void BLOCKMODE::encrypt( char file[], char outfile[], char key[] ) {
   char obuffer[BUFFSIZE];
 
   uint8_t keyblock[BKSIZE];
+  uint8_t inblock[BKSIZE];
 
   DES::sttoblk( keyblock, key );
 
@@ -46,11 +53,10 @@ void BLOCKMODE::encrypt( char file[], char outfile[], char key[] ) {
 
 
   int padding;
-  uint8_t inblock[BKSIZE];
-  printf("Got to while");
   while( (padding = 8 - infile.readsome( buffer, BUFFSIZE ) ) == 0 ){
-    printf("\n");
 
+    printf("\n");
+    memset( inblock, 0, 8 );
     DES::sttoblk( inblock, buffer );
 
     DES cipher( inblock, keyblock );
@@ -59,14 +65,13 @@ void BLOCKMODE::encrypt( char file[], char outfile[], char key[] ) {
 
     uint8_t* out = cipher.cipherText();
 
+    memset( obuffer, 0, 8 );
     DES::blktostr( out, obuffer );
 
     for (int i = 0; i < 8; i++ ) {
-      printf("%c",obuffer[i]);
-
+      printf("%d ", obuffer[i]);
     }
     ofile.write( obuffer, BUFFSIZE );
-
   }
 
   printf("\n");
@@ -78,6 +83,7 @@ void BLOCKMODE::encrypt( char file[], char outfile[], char key[] ) {
       buffer[7-i] = 0;
     }
   }
+
   DES::sttoblk( inblock, buffer );
   DES cipher( inblock, keyblock );
   cipher.encrypt();

### 2010-05-01 src/des.cpp
Optimized

@@ -3,7 +3,7 @@
 **
 **  Authors:
 **
-**     Sam Milton        (srm2997@cs.rit.edu)
+**     Sam Milton        (srm2997@rit.edu)
 **     Brian Gianforcaro (bjg1955@cs.rit.edu)
 **
 **  Description:
@@ -104,8 +104,10 @@ void DES::decrypt( void ) {
   this->algorithm( decrypt_a );
 }
 
-/* IP prime pg: 14 */
-uint8_t DES::IP[BKSIZE] = {
+
+
+
+/* Init Permuatation Definition:
   58, 50, 42, 34, 26, 18, 10, 2,
   60, 52, 44, 36, 28, 20, 12, 4,
   62, 54, 46, 38, 30, 22, 14, 6,
@@ -115,10 +117,45 @@ uint8_t DES::IP[BKSIZE] = {
   59, 51, 43, 35, 27, 19, 11, 3,
   61, 53, 45, 37, 29, 21, 13, 5,
   63, 55, 47, 39, 31, 23, 15, 7
-};
+*/
+void DES::initPermutation( uint8_t block[], uint8_t out[] ){
+  out[0]  = block[57];  out[1]  = block[49];
+  out[2]  = block[41];  out[3]  = block[33];
+  out[4]  = block[25];  out[5]  = block[17];
+  out[6]  = block[9];   out[7]  = block[1];
+  out[8]  = block[59];  out[9]  = block[51];
+  out[10] = block[43];  out[11] = block[35];
+  out[12] = block[27];  out[13] = block[19];
+  out[14] = block[11];  out[15] = block[3];
+  out[16] = block[61];  out[17] = block[53];
+  out[18] = block[45];  out[19] = block[37];
+  out[20] = block[29];  out[21] = block[21];
+  out[22] = block[13];  out[23] = block[5];
+  out[24] = block[63];  out[25] = block[55];
+  out[26] = block[47];  out[27] = block[39];
+  out[28] = block[31];  out[29] = block[23];
+  out[30] = block[15];  out[31] = block[7];
+  out[32] = block[56];  out[33] = block[48];
+  out[34] = block[40];  out[35] = block[32];
+  out[36] = block[24];  out[37] = block[16];
+  out[38] = block[8];   out[39] = block[0];
+  out[40] = block[58];  out[41] = block[50];
[…]

### 2010-05-01 src/des.cpp
blktostr clears destination

@@ -654,7 +654,7 @@ void DES::sttoblk( block_t blk, char* str ) {
 */
 
 void DES::blktostr( block_t blk, char* str ) {
-
+  memset( str, 0, 8 );
   for (int i = 0; i < 8; i++ ) {
 
     int j = i*8;

### 2010-05-01 src/test.cpp
Better test cases for block to string and viceversa.

@@ -137,26 +137,33 @@ TEST( DesBitOps, Off ) {
 
 TEST( DESutil, strtoblk ) {
 
-  /*
-  char msg[8] = "0123456";
+  char msg[9] = "01234567";
 
-  uint8_t* output1 = new uint8_t[56];
+  uint8_t expected[] = {0,0,1,1,0,0,0,0,0,0,1,1,0,0,0,1,0,0,1,1,0,0,1,0,0,0,1,1,0,0,1,1,0,0,1,1,0,1,0,0,0,0,1,1,0,1,0,1,0,0,1,1,0,1,1,0,0,0,1,1,0,1,1,1};
+
+  uint8_t* output1 = new uint8_t[64];
 
   DES::sttoblk( output1, msg );
 
-  char outmsg[8];
-  outmsg[8] = '\0';
-
-  memset(outmsg, 0, 8 );
-
-  DES::blktostr( output1, outmsg );
-
-  for (int i = 0; i < 7; i++ ) {
-    ASSERT_TRUE( outmsg[i] == msg[i] );
+  for (int i = 0; i < 8; i++ ) {
+    ASSERT_TRUE( expected[i] == output1[i] );
   }
-  */
 
 
+}
+
+TEST( DESutil, blktostr ) {
+  char expected[9] = "01234567";
+
+  uint8_t in[] = {0,0,1,1,0,0,0,0,0,0,1,1,0,0,0,1,0,0,1,1,0,0,1,0,0,0,1,1,0,0,1,1,0,0,1,1,0,1,0,0,0,0,1,1,0,1,0,1,0,0,1,1,0,1,1,0,0,0,1,1,0,1,1,1};
+
+  char outmsg[8];
+
+  DES::blktostr( in, outmsg );
+
+  for (int i = 0; i < 8; i++ ) {
+    ASSERT_TRUE( expected[i] == outmsg[i] );
+  }
 }
 
 TEST( DESAlgo, Encrypt ) {

### 2010-05-01 src/des.cpp
Small formatting changes, and removing stale func's

@@ -166,10 +166,6 @@ void DES::initPermutation( uint8_t block[], uint8_t out[] ){
   19, 13, 30,  6,
   22, 11,  4, 25
 */
-void DES::primative( uint8_t block[], uint8_t out[] ) {
-
-}
-
 
 /* Inverse Init Permutation Definition:
 
@@ -343,14 +339,14 @@ void DES::f( block_t dest, block_t R, block_t K ) {
   uint8_t rpk[48];
 
   // Expand function
-  rpk[0] = R[31] ^ K[0];   rpk[1] = R[0] ^ K[1];  
-  rpk[2] = R[1] ^ K[2];    rpk[3] = R[2] ^ K[3];  
-  rpk[4] = R[3] ^ K[4];    rpk[5] = R[4] ^ K[5];  
-  rpk[6] = R[3] ^ K[6];    rpk[7] = R[4] ^ K[7];  
-  rpk[8] = R[5] ^ K[8];    rpk[9] = R[6] ^ K[9];  
-  rpk[10] = R[7] ^ K[10];  rpk[11] = R[8] ^ K[11]; 
-  rpk[12] = R[7] ^ K[12];  rpk[13] = R[8] ^ K[13]; 
-  rpk[14] = R[9] ^ K[14];  rpk[15] = R[10] ^ K[15];
+  rpk[0]  = R[31] ^ K[0];  rpk[1]  = R[0]  ^ K[1];
+  rpk[2]  = R[1]  ^ K[2];  rpk[3]  = R[2]  ^ K[3];
+  rpk[4]  = R[3]  ^ K[4];  rpk[5]  = R[4]  ^ K[5];
+  rpk[6]  = R[3]  ^ K[6];  rpk[7]  = R[4]  ^ K[7];
+  rpk[8]  = R[5]  ^ K[8];  rpk[9]  = R[6]  ^ K[9];
+  rpk[10] = R[7]  ^ K[10]; rpk[11] = R[8]  ^ K[11];
+  rpk[12] = R[7]  ^ K[12]; rpk[13] = R[8]  ^ K[13];
+  rpk[14] = R[9]  ^ K[14]; rpk[15] = R[10] ^ K[15];
   rpk[16] = R[11] ^ K[16]; rpk[17] = R[12] ^ K[17];
   rpk[18] = R[11] ^ K[18]; rpk[19] = R[12] ^ K[19];
[…]

### 2010-05-02 src/des.cpp
Small style changes to des

@@ -29,8 +29,8 @@ using namespace std;
 **
 ** Default Constructor
 **
-** @param block, 64 bit data block made up of two 32bit int's
-** @param key, 64 bit key made up of two 32bit int's
+** @param block - 64 bit data block to encrypt.
+** @param key - 64 bit key block.
 **
 */
 
@@ -49,6 +49,7 @@ DES::DES( uint8_t* block , uint8_t* key ) {
   this->ciphertext = NULL;
   this->plaintext  = NULL;
 }
+
 
 /*
 ** Default Destructor
@@ -84,6 +85,7 @@ void DES::encrypt( void ) {
   assert( round == 0 or round >= 15 );
   this->ciphertext = new uint8_t[BKSIZE];
   this->algorithm( encrypt_a );
+  this->round = 0; /* Reset round back to one */
 }
 
 /*
@@ -102,12 +104,10 @@ void DES::decrypt( void ) {
   assert( round == 0 or round >= 15 );
   this->plaintext = new uint8_t[BKSIZE];
   this->algorithm( decrypt_a );
+  this->round = 0;
 }
 
-
-
-
-/* Init Permuatation Definition:
+/* Init Permuatationn:
   58, 50, 42, 34, 26, 18, 10, 2,
   60, 52, 44, 36, 28, 20, 12, 4,
   62, 54, 46, 38, 30, 22, 14, 6,
@@ -118,6 +118,7 @@ void DES::decrypt( void ) {
   61, 53, 45, 37, 29, 21, 13, 5,
   63, 55, 47, 39, 31, 23, 15, 7
 */
+
 void DES::initPermutation( uint8_t block[], uint8_t out[] ){
   out[0]  = block[57];  out[1]  = block[49];
   out[2]  = block[41];  out[3]  = block[33];
@@ -151,6 +152,8 @@ void DES::initPermutation( uint8_t block[], uint8_t out[] ){
   out[58] = block[46];  out[59] = block[38];
   out[60] = block[30];  out[61] = block[22];
   out[62] = block[14];  out[63] = block[6];
+
+  return;
 }
 
 
@@ -167,9 +170,7 @@ void DES::initPermutation( uint8_t block[], uint8_t out[] ){
   22, 11,  4, 25
 */
 
-/* Inverse Init Permutation Definition:
-
-*/
+/* Inverse Init Permutation Definition: */
[…]

### 2010-05-02 src/blockmode.cpp
It works.

@@ -24,6 +24,8 @@ using namespace std;
 #include "blockmode.hpp"
 #include "des.hpp"
 
+const bool DEBUG = false;
+
 BLOCKMODE::BLOCKMODE( ) {
 
 }
@@ -44,44 +46,49 @@ void BLOCKMODE::encrypt( char file[], char outfile[], char key[] ) {
 
   DES::sttoblk( keyblock, key );
 
-
+  // Files for reading / writing
   ifstream infile;
-  infile.open( file, ios::binary );
+  infile.open( file, ios::binary | ios::in | ios::ate );
 
   ofstream ofile;
-  ofile.open( outfile, ofstream::binary );
+  ofile.open( outfile, ofstream::binary | ios::out );
 
 
-  int padding;
-  while( (padding = 8 - infile.readsome( buffer, BUFFSIZE ) ) == 0 ){
+  int size = infile.tellg();   // Size of file
+  infile.seekg( 0, ios::beg );
+  int padding = 8 - (size % 8);// Amount of padding needed
+  if( padding == 0 )
+    padding = 8; // Padding cannot be 0 (pad full block)
 
-    printf("\n");
-    memset( inblock, 0, 8 );
+  // Loop through all full blocks (8 bytes) of file
+  for( int i = 0; i < size / 8; ++i ){
+    infile.read( buffer, 8 );
+
     DES::sttoblk( inblock, buffer );
 
     DES cipher( inblock, keyblock );
 
     cipher.encrypt();
 
-    uint8_t* out = cipher.cipherText();
+    DES::blktostr( cipher.cipherText(), obuffer );
 
-    memset( obuffer, 0, 8 );
-    DES::blktostr( out, obuffer );
-
-    for (int i = 0; i < 8; i++ ) {
-      printf("%d ", obuffer[i]);
+    if( DEBUG ){
+      for (int i = 0; i < 8; i++ ) {
+        printf("%02X ", obuffer[i] & 255 );
+      }
+      printf( "\n" );
     }
     ofile.write( obuffer, BUFFSIZE );
   }
 
-  printf("\n");
+  // Read remaing part of file
+  if( padding != 8 )
+    infile.read( buffer, 8 - padding );
 
-  for (int i = 0; i < padding; i++ ) {
-    if ( i+1 == padding ) {
-      buffer[7-i] = 1;
-    } else {
-      buffer[7-i] = 0;
-    }
[…]

### 2010-05-02 src/des_original.cpp
Added unoptimzed version

@@ -1,0 +1,473 @@
+/*
+**  File: des.cpp
+**
+**  Authors:
+**
+**     Sam Milton        (srm2997@rit.edu)
+**     Brian Gianforcaro (bjg1955@cs.rit.edu)
+**
+**  Description:
+**
+**     This initial implementation of the DES spec uses array's of 8-bit int's with
+**     64 elements each to represent both the DES block and DES key.
+**     Each element in the array is considered to be a single bit, and is either 0 or 1.
+*/
+
+
+#include <cstdio>
+#include <cstring>
+#include <cstdlib>
+#include <cassert>
+#include <stdint.h>
+#include <algorithm>
+
+#include "des_original.hpp"
+
+using namespace std;
+
+/*
+**
+** Default Constructor
+**
+** @param block, 64 bit data block made up of two 32bit int's
+** @param key, 64 bit key made up of two 32bit int's
+**
+*/
+
+DES::DES( block_t block , block_t key ) {
+
+  assert( block != NULL );
+  assert( key != NULL );
+
+  this->block = new uint8_t[BKSIZE];
+  this->key = new uint8_t[BKSIZE];
+
+  memcpy( this->block, block, BKSIZE );
+  memcpy( this->key, key, BKSIZE );
+  this->round = 0;
+
+  this->ciphertext = NULL;
+  this->plaintext  = NULL;
+}
+
+/*
+** Default Destructor
+*/
+
+DES::~DES( void ) {
+
+  delete[] this->block;
+  delete[] this->key;
+
+  if ( this->ciphertext != NULL ) {
+    delete[] this->ciphertext;
+  }
+
+  if ( this->plaintext != NULL ) {
+    delete[] this->plaintext;
+  }
+}
+
+/*
+** Public API to encryption algorithm of the class.
+**
+** Encipherment:
+**
+**   L[0]R[0] = IP(plain block)
+**   for 1 <= i <= 16
+**      L[i] = R[i-1]
+**      R[i] = L[i-1] xor f(R[i-1], K[i])
+**      cipher block = FP(R[16]L[16])
+*/
+
[…]

### 2010-05-02 src/des_original.hpp

@@ -1,0 +1,118 @@
+/*
+**  File: des.hpp
+*
+**  Authors: 
+** 
+**     Sam Milton        (srm2997@cs.rit.edu)
+**     Brian Gianforcaro (bjg1955@cs.rit.edu)
+**
+*/
+
+#ifndef _DES_HPP_
+#define _DES_HPP_
+
+#include <stdint.h>
+
+/* Convenience for unit testing */
+#ifdef UNITTEST
+# define protected public
+# define private   public
+#endif
+
+/* Number of rounds in cipher*/
+#define ROUNDS 16
+
+/* Initial size of block and key in bits*/
+#define BKSIZE 64
+
+/* Define which action's for the algorithm */
+enum action_t {
+  encrypt_a, decrypt_a
+};
+
+
+/* Define a block type to use */
+typedef uint8_t* block_t;
+
+class DES {
+
+  private:
+    /*
+    ** Constants as defined in the spec
+    */
+
+    /* Number of shifts for each round */
+    static uint8_t SHIFTS[ROUNDS];
+
+    /* All S-Boxes 1-8 */
+    static uint8_t SP[8][4][16];
+
+    /*
+    ** Class state
+    */
+
+    /* Data block 64 long */
+    uint8_t* block;
+
+    /* Data key block 64 long */
+    uint8_t* key;
+
+    /* Data key current round of cipher */
+    uint8_t round;
+
+    /* All scheduled key's */
+    uint8_t scheduled_keys[ROUNDS][48];
+
+    /* Encrypted cipher text */
+    uint8_t* ciphertext;
+
+    /* Decrypted plain text */
+    uint8_t* plaintext;
+
+    void mixer( uint8_t block[], uint8_t out[], uint8_t mix[], int size );
+
+    /* DES Fiestal function */
+    void f( block_t dest, block_t R, block_t K );
+
+    /* Schedule keys for all rounds */
+    void keyschedule( void );
+
+    /* Main entry point to DES cipher */
+    void algorithm( action_t action );
+
+  public:
+
+    /* Constructor */
+    DES( uint8_t* block , uint8_t* key );
+
+    /* Destructor */
+    ~DES( void );
+
+    void encrypt( void );
+
+    void decrypt( void );
+
+    uint8_t* cipherText( void ) {
+      return this->ciphertext;
+    }
+
+    uint8_t* plainText( void ) {
+      return this->plaintext;
+    }
+
+    static void sttoblk( block_t blk, char* str );
+
[…]

### 2010-05-02 src/blockmode.cpp
Fix some changes

@@ -57,11 +57,11 @@ void BLOCKMODE::encrypt( char file[], char outfile[], char key[] ) {
   int size = infile.tellg();   // Size of file
   infile.seekg( 0, ios::beg );
   int padding = 8 - (size % 8);// Amount of padding needed
-  if( padding == 0 )
+  if ( padding == 0 )
     padding = 8; // Padding cannot be 0 (pad full block)
 
   // Loop through all full blocks (8 bytes) of file
-  for( int i = 0; i < size / 8; ++i ){
+  for ( int i = 0; i < size / 8; i++ ) {
     infile.read( buffer, 8 );
 
     DES::sttoblk( inblock, buffer );
@@ -72,22 +72,23 @@ void BLOCKMODE::encrypt( char file[], char outfile[], char key[] ) {
 
     DES::blktostr( cipher.cipherText(), obuffer );
 
-    if( DEBUG ){
+    if ( DEBUG ){
       for (int i = 0; i < 8; i++ ) {
-        printf("%02X ", obuffer[i] & 255 );
+        printf( "%02X ", obuffer[i] & 255 );
       }
       printf( "\n" );
     }
+
     ofile.write( obuffer, BUFFSIZE );
   }
 
   // Read remaing part of file
-  if( padding != 8 )
+  if ( padding != 8 )
     infile.read( buffer, 8 - padding );
 
   // Pad block with a 1 followed by 0s
   buffer[8 - padding] = 1;
-  for (int i = 1; i < padding; i++ ) {
+  for ( int i = 1; i < padding; i++ ) {
       buffer[8 - i] = 0;
   }
 
@@ -96,12 +97,14 @@ void BLOCKMODE::encrypt( char file[], char outfile[], char key[] ) {
   cipher.encrypt();
   uint8_t* out = cipher.cipherText();
   DES::blktostr( out, obuffer );
-  if( DEBUG ){
+
+  if ( DEBUG ) {
     for (int i = 0; i < 8; i++ ) {
       printf("%02X ",obuffer[i] & 255 );
     }
-      printf( "\n" );
+    printf( "\n" );
   }
+
   ofile.write( obuffer, BUFFSIZE );
   infile.close();
   ofile.close();

### 2010-05-02 src/blockmode.cpp
Hook both versions into make file.

@@ -22,7 +22,12 @@
 using namespace std;
 
 #include "blockmode.hpp"
-#include "des.hpp"
+
+#ifdef DESOPTIMIZED
+  #include "des.hpp"
+#else
+  #include "des_original.hpp"
+#endif
 
 const bool DEBUG = false;
 

### 2010-05-02 paper/paper.tex
Add user manual and design section.

@@
 \usepackage{multicol}
-\usepackage{mdwlist}
-\usepackage{amssymb,amsmath}
-\usepackage{cite}
-\usepackage[bookmarks=false,colorlinks=true,linkcolor={blue},pdfstartview={XYZ null null 1.22}]{hyperref}
+
+\usepackage{acronym}
+ \usepackage[bookmarks=false,colorlinks=false,linkcolor={black},pdfstartview={XYZ null null 1.22}]{hyperref}
 \usepackage{url}
-\usepackage[T1]{fontenc}
+
+\acrodef{DES}[DES]{Data Encryption Standard}
+\acrodef{TDD}[TDD]{Test Driven Development}
+\acrodef{OOP}[OOP]{Object Oriented Programming}
 
 \begin{document}
+
 \maketitle
@@
 \begin{abstract}
-DES (Data Encryption Standard) is the 56-bit block cipher that was once used to encrypt all US Goverment non-classified documents. 
+  \ac{DES} is the 56-bit block cipher that was once used to encrypt all US Government non-classified documents.
 \end{abstract}
@@
 \newpage
-\section{The Cryptographic Primative}
+\section{The Cryptographic Primitive}
 
@@
 \section{Software Design}
-\section{Original Runtime Measurements}
-\section{Original Runtime Analysis}
+
+We chose to implement DES completely in $C++$ for this project.
+Both of the members our group were knowledgeable and comfortable
+with $C++$ and in general the $C$ world. We felt we would have a more
+generally usable and useful implementation if we went with $C++$ over
+another language (Java, Python, ... ). We attempted to take advantage ac{OOP} but
+were careful not to over do it and dive into downword spiral subclass's and
+analysing of class cohesion. The $DES$ implementation is simply a class, with actions of encrypting and
+decrypting. All state information is organized within the class. We then
+implemented a $BlockMode$ class which implements a cipher block, which is
+as you know, just a wrapper around $DES$. The user is then provided with
+two small utility functions $encrypt$ \and $decrypt$.
+
+The general ideology of this first pass is that we want the code to be clean and correct.
+We are not putting too much thought into optimization and efficiency. We are very familiar
+with the old Donal Knuth quote:
+
+   "We should forget about small efficiencies, say about 97\% of the time: premature optimization is the root of all evil"
+
+During the duration of the project we attempted to work in a \ac{TDD} model.
+All major functions and actions have associated unit-tests which we derived from various peices of
+literature on DES. These unit test's were then hooked up to run using the Google Test Framework\cite{gtest}.
+
+\section{Original Run time Measurements}
+\section{Original Run time Analysis}
 \section{Revised Software design}
-\section{Revised Runtime Measurements}
-\section{Revised Runtime Analysis}
+\section{Revised Run time Measurements}
+\section{Revised Run time Analysis}
 \section{Developers manual}
 \section{Users Manual}
-\section{ What we learned }
+
+\begin{verbatim}
+
+  Usage: ./encrypt_opt <infile> <outfile> <key>
+
+    <infile>  - The file you wish to encrypt
+    <outfile> - The resulting encrypted file
+    <key>     - 8 character string for your key
+
+  OR
+
+  Usage: ./encrypt_orig <infile> <outfile> <key>
+
+    <infile>  - The file you wish to encrypt
+    <outfile> - The resulting encrypted file
+    <key>     - 8 character string for your key
+
+\end{verbatim}
+
+\begin{verbatim}
+
+  Usage: ./decrypt_opt <infile> <outfile> <key>
+
+    <infile>  - The file you wish to decrypt
+    <outfile> - The resulting decrypted file
+    <key>     - 8 character string for your key
+
+  OR
+
+  Usage: ./decrypt_orig <infile> <outfile> <key>
+
+    <infile>  - The file you wish to decrypt
+    <outfile> - The resulting decrypted file
+    <key>     - 8 character string for your key
+
+\end{verbatim}
+
+\section{What we learned}
 \section{Future Work}
@@
 \section{References}
+
 \begin{thebibliography}{99}
-\bibitem{python}{\href{ http://python.org }{ http://python.org }}
-\bibitem{wiki} {\href{ http://en.wikipedia.org/wiki/Latitude_and_longitude_of_cities }{ http://en.wikipedia.org/wiki/Latitude\_and\_longitude\_of\_cities }}
-\bibitem{tsp} Montgomery, John {\it Tackling The Traveling Salesman Problem }{\href{ http://www.psychicorigami.com/category/tsp/ }{ http://www.psychicorigami.com/category/tsp/ } }, {\bf 2007}
-\bibitem{google} {\href{ http://code.google.com/apis/maps/documentation/staticmaps/  }{ http://code.google.com/apis/maps/documentation/staticmaps/ }}
-\bibitem{sa}{\href{ http://www.svengato.com/salesman.html }{ http://www.svengato.com/salesman.html }}
-\bibitem{dis}{\href{ http://www.geesblog.com/2009/01/calculating-distance-between-latitude-longitude-pairs-in-python/ }{ http://www.geesblog.com/2009/01/calculating-distance-between-latitude-longitude-pairs-in-python/ }}
-\bibitem{LHK}{\href{ http://www.akira.ruc.dk/~keld/research/LKH/ }{ http://www.akira.ruc.dk/~keld/research/LKH/ }}
+  \bibitem{gtest}{ \href{ http://code.google.com/p/googletest/ }{Google Unit Testing Framework}}
 \end{thebibliography}

### 2010-05-02 paper/paper.tex
Added examples and timing measurements

@@
 \usepackage{multicol}
-
-\usepackage{acronym}
+% \usepackage{acronym}
 \usepackage[bookmarks=false,colorlinks=false,linkcolor={black},pdfstartview={XYZ null null 1.22}]{hyperref}
@@
 
-\acrodef{DES}[DES]{Data Encryption Standard}
-\acrodef{TDD}[TDD]{Test Driven Development}
-\acrodef{OOP}[OOP]{Object Oriented Programming}
+% \usepackage{mdwlist}
+% \usepackage{amssymb,amsmath}
+% \usepackage{cite}
+% \usepackage[T1]{fontenc}
+
+% \acrodef{DES}[DES]{Data Encryption Standard}
+% \acrodef{TDD}[TDD]{Test Driven Development}
+% \acrodef{OOP}[OOP]{Object Oriented Programming}
 
@@
 \begin{abstract}
-  \ac{DES} is the 56-bit block cipher that was once used to encrypt all US Government non-classified documents.
+  {DES} is the 56-bit block cipher that was once used to encrypt all US Government non-classified documents.
 \end{abstract}
@@
 \section{Examples}
+
+\subsection{Example 1}
+\textit{Hello}
+\\\indent Original:
+\begin{verbatim}
+   48 65 6C 6C 6F 0A
+\end{verbatim}
+\indent Encrypted:
+\begin{verbatim}
+   94 93 81 0E 62 47 30 86
+\end{verbatim}
+
+\subsection{Example 2}
+\textit{World}
+\\\indent Original:
+\begin{verbatim}
+   57 6F 72 6C 64 0A
+\end{verbatim}
+\indent Encrypted:
+\begin{verbatim}
+   24 0B A5 B2 CB C1 F2 8B 
+\end{verbatim}
+
+\subsection{Example 3}
+\textit{Hello World}
+\\\indent Original:
+\begin{verbatim}
+   48 65 6C 6C 6F 20 57 6F
+   72 6C 64 0A
+\end{verbatim}
+\indent Encripted:
+\begin{verbatim}
+   E6 74 22 8A AC 1E 85 ED 
+   B7 3C 26 3C 4B B4 F1 77
+\end{verbatim}
+
+\subsection{Example 4}
+\textit{The quick brown fox jumps over the lazy dog}
+\\\indent Original:
+\begin{verbatim}
+   54 68 65 20 71 75 69 63 
+   6B 20 62 72 6F 77 6E 20 
+   66 6F 78 20 6A 75 6D 70 
+   73 20 6F 76 65 72 20 74 
+   68 65 20 6C 61 7A 79 20 
+   64 6F 67 0A
+\end{verbatim}
+\indent Encrypted:
+\begin{verbatim}
+   09 D1 3F 78 9D C7 66 B7 
+   46 52 A9 9F 57 92 38 65 
+   8B 40 39 1C 21 9E 77 D7 
+   9B 84 62 AF E9 22 25 1D 
+   96 61 A7 22 09 97 2C D2 
+   16 F5 D8 38 E4 FF C6 86
+\end{verbatim}
+
 \section{Software Design}
-
 We chose to implement DES completely in $C++$ for this project.
@@
 
-During the duration of the project we attempted to work in a \ac{TDD} model.
+During the duration of the project we attempted to work in a {TDD} model.
 All major functions and actions have associated unit-tests which we derived from various peices of
@@
 
-\section{Original Run time Measurements}
-\section{Original Run time Analysis}
+\section{Original Runtime Measurements}
+\paragraph{}Each test consisted of 1,000,000 encryptions of the message 123456789ABCDEF and key 133457799BBCDFF1. They were performed on a 1.8 Ghz Core2Duo computer running Ubuntu 10.04.
+\\Test 1:
+\\ \indent 10.856 seconds
+\\Test 2:
+\\ \indent 10.885 seconds
+\\Test 3:
+\\ \indent 10.912 seconds
+\\\\Average:
+\\ \indent 10.884 seconds
+\section{Original Runtime Analysis}
 \section{Revised Software design}
-\section{Revised Run time Measurements}
-\section{Revised Run time Analysis}
+\section{Revised Runtime Measurements}
+\paragraph{}Each test consisted of 1,000,000 encryptions of the message 123456789ABCDEF and key 133457799BBCDFF1. They were performed on a 1.8 Ghz Core2Duo computer running Ubuntu 10.04.
+\\Test 1:
+\\ \indent 5.294 seconds
+\\Test 2:
+\\ \indent 5.321 seconds
+\\Test 3:
+\\ \indent 5.385 seconds
+\\\\Average:
+\\ \indent 5.333 seconds
+\section{Revised Runtime Analysis}
 \section{Developers manual}

### 2010-05-02 paper/paper.tex
Added key to examples
Edited Software Design

@@
 \section{Examples}
-
+Each was encrypted using the key $24681012$.
 \subsection{Example 1}
@@
 Both of the members our group were knowledgeable and comfortable
-with $C++$ and in general the $C$ world. We felt we would have a more
+with $C++$ and in the $C$ world in general. We felt we would have a more
 generally usable and useful implementation if we went with $C++$ over
-another language (Java, Python, ... ). We attempted to take advantage ac{OOP} but
-were careful not to over do it and dive into downword spiral subclass's and
-analysing of class cohesion. The $DES$ implementation is simply a class, with actions of encrypting and
+another language (Java, Python, ... ). We attempted to take advantage of ac{OOP} but
+were careful not to over do it and dive into a downword spiral of subclasses and
+analysis of class cohesion. Our $DES$ implementation is simply a class, with actions of encrypting and
 decrypting. All state information is organized within the class. We then
-implemented a $BlockMode$ class which implements a cipher block, which is
-as you know, just a wrapper around $DES$. The user is then provided with
+implemented a $BlockMode$ class which implements a block stream cipher, which is
+just a wrapper around $DES$. The user is then provided with
 two small utility functions $encrypt$ \and $decrypt$.
 
-The general ideology of this first pass is that we want the code to be clean and correct.
+The general ideology of this first iteration is that we want the code to be clean and correct.
 We are not putting too much thought into optimization and efficiency. We are very familiar

### 2010-05-03 src/des_original.cpp
Stylistic canges and commenting

@@ -84,6 +84,7 @@ void DES::encrypt( void ) {
   assert( round == 0 or round >= 15 );
   this->ciphertext = new uint8_t[BKSIZE];
   this->algorithm( encrypt_a );
+  this->round = 0;
 }
 
 /*
@@ -102,6 +103,7 @@ void DES::decrypt( void ) {
   assert( round == 0 or round >= 15 );
   this->plaintext = new uint8_t[BKSIZE];
   this->algorithm( decrypt_a );
+  this->round = 0;
 }
 
 /* IP prime pg: 14 */
@@ -149,8 +151,8 @@ uint8_t IPP[BKSIZE] = {
 ** Mixes block into out using definitions from mix
 **
 */
-void DES::mixer( uint8_t block[], uint8_t out[], uint8_t mix[], int size ){
-  for( int i = 0; i < size; ++i ){
+void DES::mixer( uint8_t block[], uint8_t out[], uint8_t mix[], size_t size ){
+  for ( unsigned int i = 0; i < size; ++i ) {
     out[i] = block[ mix[i] - 1 ];
   }
 }
@@ -159,6 +161,8 @@ void DES::mixer( uint8_t block[], uint8_t out[], uint8_t mix[], int size ){
 /*
 **
 ** Implements the main enciphering algorithm in figure 1 pg: 13 DES spec
+**
+** @param action - should we encrypt_a or decrypt_a
 **
 */
 
@@ -218,10 +222,10 @@ uint8_t DES::SP[8][4][16] =
 {
   {
      /* S1 function pg: 19 of DES spec */
-      { 14,  4, 13,  1,  2, 15, 11,  8,  3, 10,  6, 12,  5,  9, 0,  7},
-      {  0, 15,  7,  4, 14,  2, 13,  1, 10,  6, 12, 11,  9,  5, 3,  8},
-      {  4,  1, 14,  8, 13,  6,  2, 11, 15, 12,  9,  7,  3, 10, 5,  0},
-      { 15, 12,  8,  2,  4,  9,  1,  7,  5, 11,  3, 14, 10,  0, 6, 13}
+     { 14,  4, 13,  1,  2, 15, 11,  8,  3, 10,  6, 12,  5,  9, 0,  7},
+     {  0, 15,  7,  4, 14,  2, 13,  1, 10,  6, 12, 11,  9,  5, 3,  8},
+     {  4,  1, 14,  8, 13,  6,  2, 11, 15, 12,  9,  7,  3, 10, 5,  0},
[…]

### 2010-05-03 src/des_original.hpp

@@ -69,7 +69,7 @@ class DES {
     /* Decrypted plain text */
     uint8_t* plaintext;
 
-    void mixer( uint8_t block[], uint8_t out[], uint8_t mix[], int size );
+    void mixer( uint8_t block[], uint8_t out[], uint8_t mix[], size_t size );
 
     /* DES Fiestal function */
     void f( block_t dest, block_t R, block_t K );

### 2010-05-03 src/blockmode.cpp
Style changes and spelling fix.

@@ -40,7 +40,7 @@ BLOCKMODE::~BLOCKMODE( void ) {
 }
 
 
-void BLOCKMODE::encrypt( char file[], char outfile[], char key[] ) {
+void BLOCKMODE::encrypt( char infile[], char outfile[], char key[] ) {
 
   const size_t BUFFSIZE = 8;
   char buffer[BUFFSIZE];
@@ -52,22 +52,22 @@ void BLOCKMODE::encrypt( char file[], char outfile[], char key[] ) {
   DES::sttoblk( keyblock, key );
 
   // Files for reading / writing
-  ifstream infile;
-  infile.open( file, ios::binary | ios::in | ios::ate );
-
+  ifstream ifile;
   ofstream ofile;
-  ofile.open( outfile, ofstream::binary | ios::out );
+
+  ifile.open( infile,  ios::binary | ios::in | ios::ate );
+  ofile.open( outfile, ios::binary | ios::out );
 
 
-  int size = infile.tellg();   // Size of file
-  infile.seekg( 0, ios::beg );
+  int size = ifile.tellg();   // Size of file
+  ifile.seekg( 0, ios::beg );
   int padding = 8 - (size % 8);// Amount of padding needed
   if ( padding == 0 )
     padding = 8; // Padding cannot be 0 (pad full block)
 
   // Loop through all full blocks (8 bytes) of file
   for ( int i = 0; i < size / 8; i++ ) {
-    infile.read( buffer, 8 );
+    ifile.read( buffer, 8 );
 
     DES::sttoblk( inblock, buffer );
 
@@ -87,9 +87,9 @@ void BLOCKMODE::encrypt( char file[], char outfile[], char key[] ) {
     ofile.write( obuffer, BUFFSIZE );
   }
 
-  // Read remaing part of file
+  // Read remaining part of file
   if ( padding != 8 )
-    infile.read( buffer, 8 - padding );
+    ifile.read( buffer, 8 - padding );
 
   // Pad block with a 1 followed by 0s
   buffer[8 - padding] = 1;
@@ -111,12 +111,12 @@ void BLOCKMODE::encrypt( char file[], char outfile[], char key[] ) {
   }
 
   ofile.write( obuffer, BUFFSIZE );
-  infile.close();
+  ifile.close();
   ofile.close();
 }
 
 
-void BLOCKMODE::decrypt( char file[], char outfile[], char key[] ) {
[…]

### 2010-05-05 src/blockmode.cpp
Clean up blockmode, add some macro's to make stuff clean.

@@ -18,26 +18,30 @@
 #include <stdint.h>
 #include <iostream>
 #include <fstream>
-
-using namespace std;
-
-#include "blockmode.hpp"
-
 #ifdef DESOPTIMIZED
   #include "des.hpp"
 #else
   #include "des_original.hpp"
 #endif
 
-const bool DEBUG = false;
 
-BLOCKMODE::BLOCKMODE( ) {
+#define DEBUG 0
+#include "blockmode.hpp"
 
+using namespace std;
+
+#ifdef DEBUG 
+void BLOCKMODE::printbuffer( char obuffer[] ) {
+  if ( DEBUG ) {
+    for (int i = 0; i < 8; i++ ) {
+      printf("%02X ", obuffer[i] & 255 );
+    }
+    printf( "\n" );
+  }
 }
-
-BLOCKMODE::~BLOCKMODE( void ) {
-
-}
+#else 
+ #define BLOCKMODE::printbuffer( x )     
+#endif
 
 
 void BLOCKMODE::encrypt( char infile[], char outfile[], char key[] ) {
@@ -51,7 +55,7 @@ void BLOCKMODE::encrypt( char infile[], char outfile[], char key[] ) {
 
   DES::sttoblk( keyblock, key );
 
-  // Files for reading / writing
+  /* Files for reading / writing */
   ifstream ifile;
   ofstream ofile;
 
@@ -61,12 +65,10 @@ void BLOCKMODE::encrypt( char infile[], char outfile[], char key[] ) {
 
   int size = ifile.tellg();   // Size of file
   ifile.seekg( 0, ios::beg );
-  int padding = 8 - (size % 8);// Amount of padding needed
-  if ( padding == 0 )
-    padding = 8; // Padding cannot be 0 (pad full block)
 
-  // Loop through all full blocks (8 bytes) of file
+  /* Loop through all full blocks (8 bytes) of file */
   for ( int i = 0; i < size / 8; i++ ) {
+
     ifile.read( buffer, 8 );
 
     DES::sttoblk( inblock, buffer );
@@ -77,21 +79,22 @@ void BLOCKMODE::encrypt( char infile[], char outfile[], char key[] ) {
 
     DES::blktostr( cipher.cipherText(), obuffer );
 
-    if ( DEBUG ){
-      for (int i = 0; i < 8; i++ ) {
-        printf( "%02X ", obuffer[i] & 255 );
-      }
-      printf( "\n" );
-    }
-
     ofile.write( obuffer, BUFFSIZE );
+
[…]

### 2010-05-05 src/blockmode.hpp

@@ -16,13 +16,17 @@ class BLOCKMODE {
 
   public:
 
+    /*
     BLOCKMODE( );
 
     ~BLOCKMODE( void );
+    */
 
     void encrypt( char file[], char ofile[], char key[] );
 
     void decrypt( char file[], char ofile[], char key[] );
+
+    void printbuffer( char obuffer[] );
 
 };
 

### 2010-05-05 src/test.cpp

@@ -14,6 +14,10 @@
 #include <stdint.h>
 #include <gtest/gtest.h>
 
+#ifndef  GTEST_INCLUDE_GTEST_GTEST_H_
+ #error Google testing framework required
+#endif
+
 using ::testing::InitGoogleTest;
 
 TEST( f_function, Test ){

### 2010-05-06 paper/paper.tex
Corrected timings
Add more stuff to paper.

@@
 \section{What we learned}
+
 \section{Future Work}
+
+  In the realm of future work we believe the first thing we would like to change is the cipher block mode we chose.
+  We initially chose Electronic Codebook (ECB) mode for our implementation because of the simplicity of the algorithm.
+  However for this implementation to be actually useful for legitimate security uses the program would have to be extended
+  to support alternative block modes which are not as insecure as ECB.
+
+  Another possible improvement to our implementation of DES would be to reorganize how the data input block and key block
+  are stored and used in the algorithm. Since DES has many permutations which reorganize the location of specific bits we 
+  stored the data and key blocks as array's of individual bits. It is most likely possible to keep the key and data blocks
+  organized as 8 bytes. I believe it would be possible to reduce the number of operations required total if we could bundle the 
+  bits together and then perform operations on the total bundles, much like the implementation of Skein presented in class. 
+  With this change I believe we could have a far superior implementation of DES than our current version. I am afraid that it
+  would be just out of our knowledge range to be able to actually implement it.
+
+  There also exist a few small for loops in the code which might benifit slightly from being unrolled.
+  We chose not to unroll them because in our profiling they didn't seem to be a problem. Also to unroll
+  these left over loops would affect the readability of the code tremendously, many things would need to be 
+  restructured. However unrolling all the loops might achive a slight speed up to the overall speed of implementation.
+  The difference mostly depends on the quality of the optimizations being performed by the compiler.
+
 \section{Team Member Work}
-\section{References}
 
 \begin{thebibliography}{99}
-  \bibitem{gtest}{ \href{ http://code.google.com/p/googletest/ }{Google Unit Testing Framework}}
+
+  \bibitem{ gtest }{ \href{ http://code.google.com/p/googletest/ }{ Google Unit Testing Framework } }
+  \bibitem{ spec  }{ \href{ http://csrc.nist.gov/publications/fips/fips46-3/fips46-3.pdf }{ Data Encryption Standard Specification } }
+  \bibitem{ wiki  }{ \href{ http://en.wikipedia.org/wiki/Data_Encryption_Standard }{ Wikipedia: Data Encryption Standard } }
+  \bibitem{ block }{ \href{ http://en.wikipedia.org/wiki/Block_cipher_modes_of_operation }{ Wikipedia: Block Cipher Modes } }
+
+  \bibitem{ desge }{ \href{ http://www.comms.scitech.sussex.ac.uk/fft/crypto/des_algorithm_details.txt}{ How to implement the Data Encryption Standard (DES) } }
+  \bibitem{ desg2 }{ \href{ http://orlingrabbe.com/des.htm }{ DES Illustrated } }
+  \bibitem{ gprof }{ \href{ http://linux.die.net/man/1/gprof }{ GProf Manual Page } }
+  \bibitem{ g++ }{ \href{ http://linux.die.net/man/1/g++ }{ GCC Manual Page } }
+
 \end{thebibliography}

### 2010-05-06 paper/paper.tex
Team contributions section.

@@
 \\Test 1:
-\\ \indent 10.856 seconds
+\\ \indent 10.899 seconds
 \\Test 2:
-\\ \indent 10.885 seconds
+\\ \indent 10.867 seconds
 \\Test 3:
-\\ \indent 10.912 seconds
+\\ \indent 10.870 seconds
 \\\\Average:
-\\ \indent 10.884 seconds
+\\ \indent 10.879 seconds
 \section{Original Runtime Analysis}
@@
 \\Test 1:
-\\ \indent 5.294 seconds
+\\ \indent 5.960 seconds
 \\Test 2:
-\\ \indent 5.321 seconds
+\\ \indent 6.012 seconds
 \\Test 3:
-\\ \indent 5.385 seconds
+\\ \indent 5.903 seconds
 \\\\Average:
-\\ \indent 5.333 seconds
+\\ \indent 5.958 seconds
 \section{Revised Runtime Analysis}
@@
 
-  There also exist a few small for loops in the code which might benifit slightly from being unrolled.
+  There also exist a few small for loops in the code which might benefit slightly from being unrolled.
   We chose not to unroll them because in our profiling they didn't seem to be a problem. Also to unroll
   these left over loops would affect the readability of the code tremendously, many things would need to be 
-  restructured. However unrolling all the loops might achive a slight speed up to the overall speed of implementation.
+  restructured. However unrolling all the loops might achieve a slight speed up to the overall speed of implementation.
   The difference mostly depends on the quality of the optimizations being performed by the compiler.
@@
 \section{Team Member Work}
+
+  At the start of the project we divided up th major parts of the cipher implementation for each person to do. \\
+
+  \textbf{Brian:}
+  \begin{itemize}
+    \item Unit test setup
+    \item Build system
+    \item Gathering of all constants and permutation blocks.
+    \item Key schedule 
+    \item Main DES algorithm
+  \end{itemize}
+
+  \textbf{Sam:}
+  \begin{itemize}
+    \item Bit operators and converters
+    \item Feistel function
+    \item Implementation API design
+    \item Mix function to handle mixing
+    \item Block mode implementation
+  \end{itemize}
+
+  Once these tasks were complete we worked together to integrate all parts together and debug our final product. Most of this
+  time was spent pair programming. The presentation and this paper were split up similarly but much of the work was done together, 
+  working cooperatively.
 

### 2010-05-06 paper/paper.tex
Added Original Runtime Analysis

@@
 \section{Original Runtime Analysis}
+\indent With a total run time of 10.884 seconds, this means each encryption took only .01 milliseconds.
+To find improvements, we look at our profiling.
+The most time was spent in our mixer function.
+To reduce this we decided to 'unravel' the loop.
+To do this we would have had to make a unique method for each of the different mixing procedures.
+However instead we decided to get rid of it all together by putting the unraveled contents wherever the code was called from.
+\\ \indent The next worst function was the f (feistel) function.
+One of the first things we noticed could be done for improvment was to remove the first for-loop.
+First it mixes the input, then runs through it again doing xor with the key.
+We combined the two steps into the already unraveled mix procedure.
+Also, instead of using our get method to retrieve bit values, we copied the code into the function and hard coded the values that were needed for each use.
+\\ \indent After this we analyzed the keyschedule algorithm, as this was #3 in terms of time spent executing.
+We had already internalized the mixer function.
+The innermost for loop was unraveld to remove time spent computing the the counter and comparisons.
+Also, we turned the for-loop outside of it into a do-while loop.
+This made modest savings since in cases where it only loops once, since there are no extra jumps.
+At the end of the function was one of the unraveled mix algorithms.
+We combined this with the for loop afterwards so that we weren't running through each element twice.
 \section{Revised Software design}

### 2010-05-06 paper/paper.tex
Added profiling to Original Runtime

@@
 \\ \indent 10.879 seconds
+\begin{verbatim}
+  % of   # of      # of                
+  time  Seconds    Calls    Name    
+ 40.93   18.00   51000000   DES::mixer()
+ 22.88   10.06   16000000   DES::f()
+ 20.17    8.87    1000000   DES::keyschedule()
+  8.20    3.60    1000000   DES::algorithm()
+  5.59    2.46  512000000   DES::get()
+  0.61    0.27   32000000   std::swap<T>()
+  0.53    0.23    1000000   std::swap_ranges<T,G>()
+  0.41    0.18   32000000   std::__iter_swap<T>::iter_swap<G,E>()
+  0.28    0.12   32000000   std::iter_swap<T,G>()
+  0.20    0.09              DES::on()
+  0.11    0.05    1000000   DES::DES()
+  0.02    0.01    1000000   DES::encrypt()
+  0.02    0.01    1000000   DES::~DES()
+  0.02    0.01              DES::main
+\end{verbatim}
 \section{Original Runtime Analysis}

### 2010-05-06 paper/paper.tex
Added commas to large numbers

--- 2010-05-06 paper/paper.tex
@@
 \\ \indent 10.879 seconds
+\begin{verbatim}
+  % of   # of      # of                
+  time  Seconds    Calls    Name    
+ 40.93   18.00   51,000,000   DES::mixer()
+ 22.88   10.06   16,000,000   DES::f()
+ 20.17    8.87    1,000,000   DES::keyschedule()
+  8.20    3.60    1,000,000   DES::algorithm()
+  5.59    2.46  512,000,000   DES::get()
+  0.61    0.27   32,000,000   std::swap<T>()
+  0.53    0.23    1,000,000   std::swap_ranges<T,G>()
+  0.41    0.18   32,000,000   std::__iter_swap<T>::iter_swap<G,E>()
+  0.28    0.12   32,000,000   std::iter_swap<T,G>()
+  0.20    0.09                DES::on()
+  0.11    0.05    1,000,000   DES::DES()
+  0.02    0.01    1,000,000   DES::encrypt()
+  0.02    0.01    1,000,000   DES::~DES()
+  0.02    0.01                DES::main
+\end{verbatim}
 \section{Original Runtime Analysis}

### 2010-05-06 paper/paper.tex
Formatting

@@
 
-\documentclass[12pt]{article}
+\documentclass[10pt]{article}
+
+\oddsidemargin0cm
+\topmargin-2cm     %I recommend adding these three lines to increase the 
+\textwidth16.5cm   %amount of usable space on the page (and save trees)
+\textheight23.5cm
 
@@
 Also, instead of using our get method to retrieve bit values, we copied the code into the function and hard coded the values that were needed for each use.
-\\ \indent After this we analyzed the keyschedule algorithm, as this was #3 in terms of time spent executing.
+\\ \indent After this we analyzed the keyschedule algorithm, as this was \#3 in terms of time spent executing.
 We had already internalized the mixer function.
@@
   \bibitem{ block }{ \href{ http://en.wikipedia.org/wiki/Block_cipher_modes_of_operation }{ Wikipedia: Block Cipher Modes } }
-
   \bibitem{ desge }{ \href{ http://www.comms.scitech.sussex.ac.uk/fft/crypto/des_algorithm_details.txt}{ How to implement the Data Encryption Standard (DES) } }

### 2010-05-06 paper/paper.tex
Added profiling to Optimzed Results

@@
 \\ \indent 5.958 seconds
+\begin{verbatim}
+  % of   # of      # of                  
+  Time  Seconds    Calls      Name    
+ 42.90    6.10   16,000,000   DES::f()
+ 24.40    3.47    1,000,000   DES::keyschedule()
+ 24.19    3.44    1,000,000   DES::algorithm()
+  2.04    0.29    1,000,000   DES::inverseInitPermutation()
+  2.04    0.29    1,000,000   std::swap_ranges<T,T>()
+  1.30    0.18   32,000,000   std::iter_swap<T,T>()
+  1.02    0.14   32,000,000   std::__iter_swap<G>::iter_swap<T,T>()
+  0.91    0.13    1,000,000   DES::initPermutation()
+  0.70    0.10   32,000,000   std::swap<unsigned char>()
+  0.21    0.03    1,000,000   DES::~DES()
+  0.14    0.02    1,000,000   DES::DES()
+  0.07    0.01    1,000,000   DES::encrypt()
+  0.07    0.01                main
+\end{verbatim}
 \section{Revised Runtime Analysis}

### 2010-05-06 paper/paper.tex
Formatting

--- 2010-05-06 paper/paper.tex
@@
 \topmargin-2cm     %I recommend adding these three lines to increase the 
-\textwidth16.5cm   %amount of usable space on the page (and save trees)
+\textwidth18.5cm   %amount of usable space on the page (and save trees)
 \textheight23.5cm
@@
 All major functions and actions have associated unit-tests which we derived from various peices of
-literature on DES. These unit test's were then hooked up to run using the Google Test Framework\cite{gtest}.
+literature on DES. These unit test's were then hooked up to run using the Google Test Framework \cite{gtest}.
 
@@
 \section{Developers manual}
+
+  The build environment of our software was designed to be as simple to use as possible.
+  To build just the main $encrypt$ and $decrypt$ utilties the developer must simply move to the src directory of the distribution and then issue $make$ command.
+
+{\footnotesize
+\begin{verbatim}
+$ make
+g++ -o encrypt_orig  -g3 -pg -Wall -Wextra -ansi -pedantic des_original.cpp encrypt.cpp blockmode.cpp
+g++ -o decrypt_orig  -g3 -pg -Wall -Wextra -ansi -pedantic des_original.cpp decrypt.cpp blockmode.cpp
+g++ -o encrypt_opt  -g3 -pg -Wall -Wextra -ansi -pedantic -DDESOPTIMIZED des.cpp encrypt.cpp blockmode.cpp
+g++ -o decrypt_opt  -g3 -pg -Wall -Wextra -ansi -pedantic -DDESOPTIMIZED des.cpp decrypt.cpp blockmode.cpp
+\end{verbatim}
+}
+
+  This will build 4 binaries in total, an optimized $encrypt$ and $decrypt$ as well as the original $encrypt$ and $decrypt$.
+  The distribution also comes with unit-tests and a sample program to perform performance tests with.
+  In order to build the unit tests you will need the Google Testing Frame work available from http://code.google.com/p/googletest/ or if you are
+  on a Ubuntu/Debian machine:
+
+{\footnotesize
+\begin{verbatim}
+$ sudo apt-get install libgtest-dev
+\end{verbatim}
+}
+
+You can then build the testers:
+
+{\footnotesize
+\begin{verbatim}
+$ make tester
+g++ -o tester_orig  -lgtest -g3 -pg -Wall -Wextra -ansi -pedantic -DUNITTEST blockmode.cpp des_original.cpp test.cpp
+g++ -o perftest_orig  -g3 -pg -Wall -Wextra -ansi -pedantic  -DUNITTEST des_original.cpp performance.cpp
+g++ -o tester_opt  -lgtest -g3 -pg -Wall -Wextra -ansi -pedantic -DUNITTEST -DDESOPTIMIZED blockmode.cpp des.cpp test.cpp
+g++ -o perftest_opt  -g3 -pg -Wall -Wextra -ansi -pedantic  -DUNITTEST -DDESOPTIMIZED des.cpp performance.cpp
+\end{verbatim}
+}
+
+Finally you can run the testers:
+
+{\footnotesize
+\begin{verbatim}
+$ make test
+\end{verbatim}
+}
+
 \section{Users Manual}
 
+In order to encrypt a file using out implementation the user must run the executable \textbf{encrypt\_opt}.
 \begin{verbatim}
@@
 
-  OR
-
-  Usage: ./encrypt_orig <infile> <outfile> <key>
-
-    <infile>  - The file you wish to encrypt
-    <outfile> - The resulting encrypted file
-    <key>     - 8 character string for your key
-
 \end{verbatim}
 
+The user has a similar choice for decrypting, both an optimized and original implementation exist.
 \begin{verbatim}
@@
     <key>     - 8 character string for your key
-
-  OR
-
-  Usage: ./decrypt_orig <infile> <outfile> <key>
-
-    <infile>  - The file you wish to decrypt
-    <outfile> - The resulting decrypted file
     <key>     - 8 character string for your key

### 2010-05-06 paper/paper.tex
Formatting developers manual.

@@
 We combined this with the for loop afterwards so that we weren't running through each element twice.
+
 \section{Revised Software design}
+
 \section{Revised Runtime Measurements}
@@
 \end{verbatim}
+
 \section{Revised Runtime Analysis}
+
 \section{Developers manual}
@@
 
-You can then build the testers:
+You can then build the testers and run them all automatically:
 
@@
 \begin{verbatim}
-$ make tester
+$ make test
 g++ -o tester_orig  -lgtest -g3 -pg -Wall -Wextra -ansi -pedantic -DUNITTEST blockmode.cpp des_original.cpp test.cpp
@@
 g++ -o perftest_opt  -g3 -pg -Wall -Wextra -ansi -pedantic  -DUNITTEST -DDESOPTIMIZED des.cpp performance.cpp
-\end{verbatim}
-}
 
-Finally you can run the testers:
-
-{\footnotesize
-\begin{verbatim}
-$ make test
+.... test output ...
 \end{verbatim}

### 2010-05-06 paper/paper.tex
Started Revised Runtime Analysis

@@
 \section{Revised Runtime Analysis}
-
+  One of the major changes was the f (feistel) function.
+It went from 23\% to 43\% of the total runtime.
+This change is largely because we removed the mixer function.
+Mixer accounted for 41\% of the runtime, 63\% of mixer's calls were from feistel.
+From this we can deduce that 26\% out of the 41\% was directly due to f, and we could say that f really caused 23\% + 26\% = 49\% of the runtime.
+From this we can see that the f function's overall contribution to execution time went down without removing any of its duties.
 \section{Developers manual}

### 2010-05-06 paper/paper.tex
Little changes to paper.

--- 2010-05-06 paper/paper.tex
@@
 \topmargin-2cm     %I recommend adding these three lines to increase the
-\textwidth18.5cm   %amount of usable space on the page (and save trees)
+\textwidth16.5cm   %amount of usable space on the page (and save trees)
 \textheight23.5cm
@@
 \newpage
-\section{The Cryptographic Primitive}
+
+\part{The Cryptographic Primitive}
 
@@
 
-  The build environment of our software was designed to be as simple to use as possible.
+\indent  The build environment of our software was designed to be as simple to use as possible.
+  The only dependency to compiler our encrypt and decrypt programs is any modern C++ compiler, we recommend the GNU Compiler Collection's G++.
   To build just the main $encrypt$ and $decrypt$ utilties the developer must simply move to the src directory of the distribution and then issue $make$ command.
@@
 \begin{verbatim}
+
 $ make
@@
 g++ -o decrypt_opt  -g3 -pg -Wall -Wextra -ansi -pedantic -DDESOPTIMIZED des.cpp decrypt.cpp blockmode.cpp
+
 \end{verbatim}
 }
-
-  This will build 4 binaries in total, an optimized $encrypt$ and $decrypt$ as well as the original $encrypt$ and $decrypt$.
-  The distribution also comes with unit-tests and a sample program to perform performance tests with.
-  In order to build the unit tests you will need the Google Testing Frame work available from http://code.google.com/p/googletest/ or if you are
-  on a Ubuntu/Debian machine:
-
+\noindent  This will build 4 binaries in total, an optimized $encrypt$ and $decrypt$ as well as our original implementation $encrypt$ and $decrypt$.
+  The two set's of binaries are denoted by the "\_opt" for optimized and "\_orig" for original appended to each.
+  Our distribution also comes with unit-tests and a sample program to perform performance tests with.
+  In order to build the unit tests you will need the Google Testing Framework available from http://code.google.com/p/googletest/ or it is
+  probabably available in your package manager if you are using a unix system. For example if you are on any Ubuntu/Debian machine you should just need to run the following command:
 {\footnotesize
@@
 }
-
-You can then build the testers and run them all automatically:
-
+\noindent You can then build the testers and run them all automatically:
 {\footnotesize

### 2010-05-06 paper/paper.tex
Expanded Software Design section

@@
 \section{Software Design}
+Our implementation of DES was done entirely in C++. This was chosen since both members were knowledgeable and comfortable using C++.
+
+Our DES implementation is contained in one class.
+The implementation is spread across 4 functions; algorithm(), keyschedule(), mixer(), f().
+Algorithm() is the starting point.
+It calls keyschedule() to set up the keys, uses mixer() to perform the initial permutation, then runs the 16 rounds, and ends by using mixer() again to do the inverse initial permutation.
+Each round first picks the round's key, calls the f() function, and then swaps the two halves of the key.
+
+The keyschedule() function first uses mixer() to do permutation choice 1.
+It then generates 16 sub keys.
+Each sub key is generated by shifting the bits 1 or 2 times then using mixer() to do permutation choice 2.
+
+The mixer() function isn't specifically part of the DES spec.
+DES has many instances where bits must be reordered, we wrote one function, mixer(), which can be passed data and an array defining how to rearrange things and it will rearrange them.
+
+The last function, f(), also known as the Feistel function, is the heart of the encryption.
+It first uses mixer() to do the expansion permutation, then xor's the result with the sub key.
+This output is split into 8-bit block and each block is sent through an S-Box.
+The output from the S-Boxes is then concatenated run through mixer() using the primitive permutation.
+\\ \\ \\
 We chose to implement DES completely in $C++$ for this project.

### 2010-05-06 paper/paper.tex
Added What we Learned content

@@
 \section{What we learned}
+One of the very first things we learned is that in order to implement the specification, we would have to put much thought and careful interpretation into it.
+A big contributing factor to this is that the documentation for cryptographic specs is more complicated that what we're used to.
 
+We quickly discovered after our very first implementation, that even after trying to carefully follow the spec, it is very easy to make a mistake which messes everything up.
+We spent a fair amount of time debugging our algorithm to get it right.
+
+It was also surprising to see how much performance was regained simply by unrolling loops and removing function calls. We didn't expect to see as much of an improvement from that as we got.
 \section{Future Work}

### 2010-05-06 paper/paper.tex
Combined Orig Runtime Analysis with Design Changes

@@
 \end{verbatim}
-\section{Original Runtime Analysis}
+\section{Original Runtime Analysis and Design Changes}
 \indent With a total run time of 10.884 seconds, this means each encryption took only .01 milliseconds.
@@
 To do this we would have had to make a unique method for each of the different mixing procedures.
-However instead we decided to get rid of it all together by putting the unraveled contents wherever the code was called from.
+However instead we decided to get rid of it all together by putting the unraveled contents wherever mixer() was called from.
 \\ \indent The next worst function was the f (feistel) function.
@@
 First it mixes the input, then runs through it again doing xor with the key.
-We combined the two steps into the already unraveled mix procedure.
+We combined the two steps into the already unraveled and inlined mix procedure.
 Also, instead of using our get method to retrieve bit values, we copied the code into the function and hard coded the values that were needed for each use.
@@
 We combined this with the for loop afterwards so that we weren't running through each element twice.
-
-\section{Revised Software design}
 

### 2010-05-06 paper/paper.tex
Add more stuff

@@
 \usepackage{graphicx}
+\usepackage{color}
 \usepackage{multicol}
@@
 
-\begin{abstract}
-  {DES} is the 56-bit block cipher that was once used to encrypt all US Government non-classified documents.
-\end{abstract}
+%\begin{abstract}
+%  {DES} is the 56-bit block cipher that was once used to encrypt all US Government non-classified documents.
+%$\end{abstract}
 
@@
 
-\part{The Cryptographic Primitive}
+\columnseprule .4pt
+
+\definecolor{light-gray}{gray}{0.95}
+
+\def \columnseprulecolor{ \color{light-gray} }
+
+\section{The Cryptographic Primitive}
+
+  {DES} is the 64-bit block cipher that was once used to encrypt all US Government non-classified documents.
+  It was developed as a collaboration between NSA and IBM in the 1970's and was eventually selected
+  by the National Institute of Standards and Technologies and was  standardised on January 15, 1977 as FIPS-46.
+  As mentioned before the cipher has a 64-bit data block and uses a 64-bit key block, it should be noted though that
+  only 56 bits of the key are ever actually used. The algorithm is based around sixteen rounds each with it's own
+  shortened subkey (48 bits) based on the original key block. At the beginning of the cipher the data bock is permuted
+  and is then inversely permuted
 
@@
 Each was encrypted using the key $24681012$.
+
+\begin{multicols}{3}
+
+\hbadness=10000
+
   \subsection{Example 1}
@@
   \end{verbatim}
+
+\end{multicols}
 
@@
 The output from the S-Boxes is then concatenated run through mixer() using the primitive permutation.
-\\ \\ \\
-We chose to implement DES completely in $C++$ for this project.
-Both of the members our group were knowledgeable and comfortable
-with $C++$ and in the $C$ world in general. We felt we would have a more
-generally usable and useful implementation if we went with $C++$ over
-another language (Java, Python, ... ). We attempted to take advantage of ac{OOP} but
-were careful not to over do it and dive into a downword spiral of subclasses and
-analysis of class cohesion. Our $DES$ implementation is simply a class, with actions of encrypting and
-decrypting. All state information is organized within the class. We then
-implemented a $BlockMode$ class which implements a block stream cipher, which is
-just a wrapper around $DES$. The user is then provided with
-two small utility functions $encrypt$ \and $decrypt$.
 
@@
 
+\begin{quote}
    "We should forget about small efficiencies, say about 97\% of the time: premature optimization is the root of all evil"
+\end{quote}
 
@@
 
-.... test output ...
+... test output ...
 \end{verbatim}
@@
 
-In order to encrypt a file using out implementation the user must run the executable \textbf{encrypt\_opt}.
+\noindent In order to encrypt a file using our implementation the user must run the executable \textbf{encrypt\_opt}.
 \begin{verbatim}
@@
 
-The user has a similar choice for decrypting, both an optimized and original implementation exist.
+\noindent The user has a similar choice for decrypting, both an optimized and original implementation exist.
 \begin{verbatim}
@@
 
-  \bibitem{ gtest }{ \href{ http://code.google.com/p/googletest/ }{ Google Unit Testing Framework } }
+  \bibitem{gtest}{ \href{ http://code.google.com/p/googletest/ }{ Google Unit Testing Framework } }
   \bibitem{ spec  }{ \href{ http://csrc.nist.gov/publications/fips/fips46-3/fips46-3.pdf }{ Data Encryption Standard Specification } }

### 2010-05-06 paper/paper.tex
Modify color
Paper finalizations.

@@
-\title{ Data Encryption Standard Implementation 4003-482-01 RIT }
+\title{ Data Encryption Standard Implementation 4003-482-01 RIT \\ Team BitShift }
 
@@
 
-\definecolor{light-gray}{gray}{0.95}
+% \raggedcolumns
+
+\definecolor{light-gray}{gray}{0.50}
 
@@
 
-  {DES} is the 64-bit block cipher that was once used to encrypt all US Government non-classified documents.
-  It was developed as a collaboration between NSA and IBM in the 1970's and was eventually selected
-  by the National Institute of Standards and Technologies and was  standardised on January 15, 1977 as FIPS-46.
-  As mentioned before the cipher has a 64-bit data block and uses a 64-bit key block, it should be noted though that
-  only 56 bits of the key are ever actually used. The algorithm is based around sixteen rounds each with it's own
-  shortened subkey (48 bits) based on the original key block. At the beginning of the cipher the data bock is permuted
-  and is then inversely permuted
+\indent DES\cite{spec} is the 64-bit block cipher that was once used to encrypt all US Government 
+  non-classified documents.  It was developed primarily by IBM with collaboration with the NSA
+  in the 1970's and was eventually selected by the National Institute of Standards   
+  and Technologies (NIST) as the first standarized block cipher.
+  It was standardised on January 15, 1977 as FIPS-46. As mentioned 
+  before the cipher has a 64-bit data block and uses a 64-bit key block, it should be 
+  noted though that only 56 bits of the key are ever actually used. The algorithm is 
+  based around sixteen rounds each with it's own shortened subkey (48 bits) based on 
+  the original key block. DES has since been broken, one attack used a large distributed computing network
+  to crack a DES key in a mere 22 hours of compute time \cite{deepcrack}. Many cryptographers cite
+  it's primarily weekness of DES is it's 56-bit key, being far too small. 
+  As of July 2005 NIST has withdrawn DES (FIPS-46) as a standardized
+  cryptographic primative for govenment use\cite{withdrawn}. The cipher served the the US government
+  and the world at large for around twenty eight years of secure service.
 
@@
   \subsection{Example 1}
-  \textit{Hello}
-  \\\indent Original:
+
+  \textit{Hello} \\
+\noindent Original:
   \begin{verbatim}
@@
   \subsection{Example 2}
-  \textit{World}
-  \\\indent Original:
+
+  \textit{World} \\
+\noindent Original:
   \begin{verbatim}
@@
   \subsection{Example 3}
-  \textit{Hello World}
-  \\\indent Original:
+
+  \textit{Hello World} \\
+\noindent Original:
   \begin{verbatim}
@@
   \end{verbatim}
-  \indent Encripted:
+  \indent Encrypted:
   \begin{verbatim}
@@
 \subsection{Example 4}
-\textit{The quick brown fox jumps over the lazy dog}
-\\\indent Original:
+
+\textit{The quick brown fox jumps over the lazy dog} \\
+
+\noindent Original:
 \begin{verbatim}
@@
 
+During the duration of the project we attempted to work in a Test Driven Development (TDD) model.
+All major functions and actions have associated unit-tests which we derived from various peices of
+literature on DES. These unit test's were then hooked up to run using the Google Test Framework \cite{gtest}.
 The general ideology of this first iteration is that we want the code to be clean and correct.
@@
 
+{
+\textbf{
+\large{
 \begin{quote}
-   "We should forget about small efficiencies, say about 97\% of the time: premature optimization is the root of all evil"
+   ``We should forget about small efficiencies, say about 97\% of the time: premature optimization is the root of all evil.''
 \end{quote}
-
-During the duration of the project we attempted to work in a {TDD} model.
-All major functions and actions have associated unit-tests which we derived from various peices of
-literature on DES. These unit test's were then hooked up to run using the Google Test Framework \cite{gtest}.
+}
+}
+}
 
@@
 \paragraph{}Each test consisted of 1,000,000 encryptions of the message 123456789ABCDEF and key 133457799BBCDFF1. They were performed on a 1.8 Ghz Core2Duo computer running Ubuntu 10.04.
-\\Test 1:
-\\ \indent 10.899 seconds
-\\Test 2:
-\\ \indent 10.867 seconds
-\\Test 3:
-\\ \indent 10.870 seconds
-\\\\Average:
-\\ \indent 10.879 seconds
+
+\begin{multicols}{4}
+Test 1: 10.899 sec
+
+Test 2: 10.867 sec
+
+Test 3: 10.870 sec
+
+Average: $ 10.879 $ sec
+\end{multicols}
 \begin{verbatim}
@@
 \paragraph{}Each test consisted of 1,000,000 encryptions of the message 123456789ABCDEF and key 133457799BBCDFF1. They were performed on a 1.8 Ghz Core2Duo computer running Ubuntu 10.04.
-\\Test 1:
-\\ \indent 5.960 seconds
-\\Test 2:
-\\ \indent 6.012 seconds
-\\Test 3:
-\\ \indent 5.903 seconds
-\\\\Average:
-\\ \indent 5.958 seconds
+\begin{multicols}{4}
+Test 1:  5.960 sec
+
+Test 2:  6.012 sec
+
+Test 3:  5.903 sec
+
+Average: 5.958 sec
+\end{multicols}
 \begin{verbatim}
@@
 From this we can see that the f function's overall contribution to execution time went down without removing any of its duties.
+
+\pagebreak
 \section{Developers manual}
@@
 }
-
 \section{Users Manual}
@@
 \begin{verbatim}
-
   Usage: ./encrypt_opt <infile> <outfile> <key>
@@
     <key>     - 8 character string for your key
-
 \end{verbatim}
 
-\noindent The user has a similar choice for decrypting, both an optimized and original implementation exist.
+\noindent \textbf{Ex: ./encrypt\_opt file.jpg file.enc.jpg 12345678} \\
+
+\noindent Similarly to decrypt the file  user has a similar choice for decrypting, both an optimized and original implementation exist.
 \begin{verbatim}
-
   Usage: ./decrypt_opt <infile> <outfile> <key>
@@
     <key>     - 8 character string for your key
+\end{verbatim}
 
-\end{verbatim}
+\noindent \textbf{Ex: ./decrypt\_opt file.enc.jpg file.dec.jpg 12345678} \\
+
+The user should be aware that two set's of of the program exist and are denoted by the \textbf{\_opt} for optimized and \textbf{\_orig} for original appended to each.
+These exist so you can easily see the performance difference between the two implementations your self on large files.
 
@@
 
-  \textbf{Brian:}
+  \noindent \textbf{Brian:}
   \begin{itemize}
@@
 
-  \textbf{Sam:}
+ \noindent \textbf{Sam:}
   \begin{itemize}
@@
 
+  \bibitem{spec}{ \href{ http://csrc.nist.gov/publications/fips/fips46-3/fips46-3.pdf }{ Data Encryption Standard Specification } }
+  \bibitem{deepcrack}{ \href{ http://en.wikipedia.org/wiki/EFF_DES_cracker }{  DES Challenge  } }
+  \bibitem{withdrawn}{ \href{ http://csrc.nist.gov/publications/fips/05-9945-DES-Withdrawl.pdf }{ Federal Register document, stating the withdrawl of FIPS-46 } }
   \bibitem{gtest}{ \href{ http://code.google.com/p/googletest/ }{ Google Unit Testing Framework } }
-  \bibitem{ spec  }{ \href{ http://csrc.nist.gov/publications/fips/fips46-3/fips46-3.pdf }{ Data Encryption Standard Specification } }
   \bibitem{ wiki  }{ \href{ http://en.wikipedia.org/wiki/Data_Encryption_Standard }{ Wikipedia: Data Encryption Standard } }

### 2010-05-06 paper/paper.tex
Noted compiler optimization levels

@@
 \section{Original Runtime Measurements}
-\paragraph{}Each test consisted of 1,000,000 encryptions of the message 123456789ABCDEF and key 133457799BBCDFF1. They were performed on a 1.8 Ghz Core2Duo computer running Ubuntu 10.04.
+\paragraph{}Each test consisted of 1,000,000 encryptions of the message 12 34 56 78 9A BC DE F and key 13 34 57 79 9B BC DF F1.
+They were performed on a 1.8 Ghz Core2Duo computer running Ubuntu 10.04.
+The times were made with level 2 compiler optimiz
+ation. Our profiling data had all compiler optimization turned off.
 
@@
 \section{Revised Runtime Measurements}
-\paragraph{}Each test consisted of 1,000,000 encryptions of the message 123456789ABCDEF and key 133457799BBCDFF1. They were performed on a 1.8 Ghz Core2Duo computer running Ubuntu 10.04.
+\paragraph{}Each test consisted of 1,000,000 encryptions of the message 12 34 56 78 9A BC DE F and key 13 34 57 79 9B BC DF F1.
+They were performed on a 1.8 Ghz Core2Duo computer running Ubuntu 10.04.
+The times were made with level 2 compiler optimization.
+Our profiling data had all compiler optimization turned off.
 \begin{multicols}{4}
